---
title: 'PA_1 template Coursera Reproducible Research Project #1'
author: "Edward S."
date: "December 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Set Up

```{r Set up}
library(Hmisc)
```




# 1A: Code for Reading in the Activity Dataset:

```{r Read in the Data}

if(!file.exists("C:/Users/edwsp/Desktop/repdata%2Fdata%2Factivityactivity.csv")){
    fileUrl<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    download.file(fileUrl, 
    destfile = paste0(getwd(),'/repdata%2Fdata%2Factivity.zip'))
}

if(!file.exists("activity.csv")){
      unzip("C:/Users/edwsp/Desktop/repdata%2Fdata%2Factivity/activity.csv")}

A <- read.csv("data/activity.csv", header=TRUE)
```

#1A: ALternative Read In of Data:

```{r Alternative Read In of Data}

if(!file.exists("(paste0(getwd(),'/repdata%2Fdata%2Factivity.zip')")) {
        temp <- tempfile()
        download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",
                      temp)
        unzip(temp)
        unlink(temp)
}

A<-read.csv("activity.csv")
```



#1B: Code for Processing the Data:

```{r Process the Data, results="hide"}
str(A)
attach(A)
```



#2: Determine the number of Steps per day and Histogram of Total Number of Steps per Day:

```{r Hist of Steps}
SPD<-tapply(steps, date, sum)
hist(SPD, main="Steps per Day", col="blue", xlab="Steps per Day")
```


#3. Mean and Median Number of Steps per Day:

```{r Mean and Median Steps}
summary(SPD)
```


#4. Time Series Plot of Average Number of Steps Taken:

```{r Time Series Plot}
SPI.t<-tapply(steps, interval, mean, na.rm=TRUE)
summary(SPI.t)
plot(as.numeric(names(SPI.t)), SPI.t, type="l", main = "Pattern of Steps per Interval", xlab="Intervals", ylab="Steps", col="blue")
```



#5. The 5 Minute Interval that on Average has the Maximum Number of Steps:

```{r Max 5}
Max.steps<-names(sort(SPI.t, decreasing = TRUE)[1])
Max.steps
```



#6. Code to Describe and Show Strategy to Impute Missing Data:
I'm simple- I'm just going to impute median number of steps for each interval when there's a missing value.

```{r Imputation, results="hide"}
sum(is.na(steps))
(sum(is.na(steps)))/(length(A$steps))

SPI.a<-aggregate(steps, list(interval), mean, na.rm=TRUE)
colnames(SPI.a)<-c("interval", "SPI")
B<-merge(A, SPI.a, by="interval", all.x=TRUE)
B$steps.i<-ifelse(is.na(B$steps), B$SPI, B$steps)
attach(B)
```



#7. Histogram of the Total Number of Steps each Day after Imputation:

```{r Hist after Imputed}
SPD.i<-tapply(steps.i, date, sum)

hist(SPD.i, main="Steps per Day with and without Imputation", col=rgb(1,0,0,0.5), xlab="Steps per Day") 

legend("topright", c("Imputed", "Non-imputed", "Overlap"), col=c("red", "blue", "purple"), lwd=10)

hist(SPD, main="Steps per Day", col=rgb(0,0,1,0.5), add=TRUE)

SPD.i.mean<-mean(SPD.i)
SPD.i.median<-median(SPD.i)

SPD.mean<-mean(SPD, na.rm=TRUE)
SPD.median<-median(SPD, na.rm=TRUE)

Mean.dif<-SPD.i.mean-SPD.mean
Median.dif<-SPD.i.median-SPD.median

SPD.i.mean
SPD.i.median
SPD.mean
SPD.median
Mean.dif
Median.dif
```



#8. Panel Plot comparing Average Number of Steps taken per 5 Min Interval across Weekends and Weekdays

```{r Panel Plot}
library(lattice)

B$Day<-weekdays(as.Date(B$date))
B$Daytype<-ifelse(((B$Day=="Saturday")|(B$Day=="Sunday")), "Weekend", "Weekday")


SPI.i<-aggregate(B$steps.i, list(B$interval), mean, na.rm=TRUE)
colnames(SPI.i)<-c("interval", "SPI.i")
C<-merge(B, SPI.i, by="interval", all.x=TRUE)

SPI.i.Wkday<-C[which(C$Daytype=="Weekday"),]
SPI.i.Wkend<-C[which(C$Daytype=="Weekend"),]

par(mfrow=c(1,2))
    
plot(SPI.i.Wkday$interval, SPI.i.Wkday$SPI.i,
     xlab = "Interval", 
     ylab = "Steps Imputed", 
     main = "Pattern of Steps per Interval (Weekdays)", 
     type = "l",
     col="blue")

plot(SPI.i.Wkend$interval, SPI.i.Wkend$SPI.i, 
     xlab = "Interval", 
     ylab = "Steps Imputed", 
     main = "Pattern of Steps per Interval (Weekends)", 
     type = "l",
     col="red")
```



#9. All of the R Code needed to Reproduce the Plots can be found in the above report.

